\documentclass[11pt]{exam}
\usepackage{../commonheader}

\title{Syllabus \& Linear Systems}
\date{Week 1, Session 1}

\begin{document}
\maketitle

\section{SI Logistics}
    \begin{itemize}
        \item By now, you should hopefully be in the \href{https://iu.instructure.com/courses/2278911}{Math 303 SI companion course}.
        If you're not, please let me know via email or Canvas message!
        \item Fill out the \href{https://www.when2meet.com/?25969358-aRu9i}{when2meet survey} to let me know when you're free for
        SI sessions and office hours beginning next week!
        \item Here's \href{https://asciimath.org/}{a guide to AsciiMath syntax} that you may want to bookmark. To enter ``math mode``
        in a MyOpenMath window (e.g. on quizzes), just surround some math text with backticks. For example,
        \verb|`[[a,b],[c,d]]`| generates a matrix $\begin{bmatrix} a & b \\ c & d \end{bmatrix}$.
        \item To my knowledge, discussion posts use Latex instead of AsciiMath, so the above should hopefully only be relevant to quizzes/exams. 
    \end{itemize}

\pagebreak
\section{Opener/Intro Activity}
    Before we start the session, let's get to know each other! As we go around the room, share some or all of the following:
    \begin{enumerate}
        \item Name, where you're from
        \item Major, how long you've been at IU, and your current career/academic goal
        \item Your favorite class you've taken up to this point
        \item Would you rather live in the city, suburbs, or country? How come?
        \item What's a favorite song or artist you've discovered recently?
    \end{enumerate}

\pagebreak
\section{Syllabus Overview}
    \vspace{10px}
    \subsection{Homework}
    \begin{itemize}
        \item For homeworks, you can retry problems twice with the answer being shown on your third attempt.
        \item If you're unable to get the problem correct even with your retries, you can select
        "Try a similar problem" to get a new version of the problem to solve. If you're able to complete this new problem,
        you will still get full credit! This means you can get full credit on every homework as long as you put in the work.
        \item Homework is due every Saturday night at 11:59 PM Eastern.
    \end{itemize}

    \vspace{10px}
    \subsection{Quizzes}
    \begin{itemize}
        \item Quizzes require submitting an answer and an explanation. 50\% credit is given for the correct answer, 50\% for the explanation.
        \item There will be a box under each quiz problem where you can type your explanation using \href{https://asciimath.org/}{AsciiMath}.
        \item For the accuracy part, you will generally have 3 attempts per question with a 30\% penalty for each incorrect attempt.
        \item Quizzes have a 3-hour time limit beginning when you start them, and are due every Monday night at 11:59 PM Eastern.
        \item \textbf{TIP:} You generally have plenty of time on quizzes. Use the extra time to check your work before submitting to make sure
        you avoid careless mistakes that cost you 30\% each! 
    \end{itemize}
    
    \vspace{10px}
    \subsection{Exams}
    \begin{itemize}
        \item There will be two exams, a midterm and a final. Both will have a 4-hour time limit and be due on Fridays at 11:59 PM Eastern.
        \item There won't be quizzes on exam weeks.
        \item The exams will be identical to quizzes in format and grading.
        \item Prior to each exam, there will be a special exam review assignment on top of the normal work for the week. This is \textit{graded} and
        generally very helpful to prepare for the exams, so make sure to do them!
    \end{itemize}
    
    \vspace{10px}
    \subsection{Discussion Boards}
    \begin{itemize}
        \item To get full credit on discussion boards, you must contribute (post or reply meaningfully) at least 5 times each week.
        \item The "Posts on Discussion Board" assignment itself is ungraded; it's just there to provide the necessary info for that week's discussion.
        \item Posts 1 and 2 will be solving your two problems from the "Posts on Discussion Board" assignment for the week. These problems will be 
        different for everyone!
        \item Post 3 will consist of finding a video related to the week's topics on YouTube and embedding it in a discussion post. Make sure no one else
        has already chosen your video (you may not get credit for this post if so)!
        \item Posts 4 and 5 should be meaningful responses to a classmate's post. Some ideas for meaningful replies are correcting a mistake in their
        solution, offering an alternative solution method, or expanding on a topic covered by their problem.
    \end{itemize}

\pagebreak
\section{System of Equations, Geometry}
    \begin{questions}
        \item Use the \href{https://www.desmos.com/calculator}{Desmos Graphing Calculator} to plot the following two-variable linear systems, and describe
        their solutions.
        \begin{enumerate}
            \item $\begin{cases}
                x + y = 3 \\
                y - x = 5
            \end{cases}$
            \item $\begin{cases}
                y = 3x - 5 \\ 
                3y -9x = - 5
            \end{cases}$
            \item $\begin{cases}
                x + y = 3 \\
                y - x = 5 \\
                x = -1 \\
                y + 4x = 0
            \end{cases}$
            \item $\begin{cases}
                y - 2x = -4 \\
                y = x - 2 \\
                2y = 2x - 4 \\
                x + y = 1
            \end{cases}$
        \end{enumerate}

        \item Suppose you have a system of 4 equations in two variables. For each number of solutions below, describe or draw how a system with that
        many solutions may look geometrically:
        \begin{enumerate}
            \item No solution
            \item A unique solution
            \item An infinite number of solutions
        \end{enumerate}
    \end{questions}

\pagebreak
\section{System of Equations, Algebraic Procedures}
    \begin{questions}
        \item Recall the definitions of homogeneous systems, consistent systems, and inconsistent systems. Determine whether the following statements
        are true or false and explain why.
        \begin{enumerate}
            \item Homogeneous linear systems are always consistent.
            \item A system of linear equations can have exactly three solutions.
            \item A system of linear equations can have a solution which does not satisfy all conditions of the system.
            \item A system of more than two linear equations in two variables cannot be consistent.
            \item A two-variable homogeneous system may have solutions besides $(0,0)$.
        \end{enumerate}

        \vspace{20px}
        \item Use elementary row operations and back substitution to solve the following systems (or determine it has no solutions):
        \begin{enumerate}
            \item $\begin{cases}
                3x + 2y = 6 \\
                2x + 3y = 9
            \end{cases}$
            \item $\begin{cases}
                4x - 3y = 10 \\
                \frac{1}{2}x - y = 0
            \end{cases}$
            \item $\begin{cases}
                3x + y - z = 9 \\
                2x - 2y + z = -3 \\
                x + y + z = 7
            \end{cases}$
            \item $\begin{cases}
                4x - 3y = 10 \\
                8x - 6y = 20
            \end{cases}$
            \item $\begin{cases}
                6x - 3y = 2 \\
                3x - \frac{3}{2}y = 3
            \end{cases}$
            \item $\begin{cases}
                12x - 9y = 3 \\
                12x + y = 0 \\
                x + y = 5
            \end{cases}$
        \end{enumerate}
    \end{questions}

\pagebreak
\section{Closing - Informal Quiz}
    \begin{questions}
        \item Determine whether the following statements are true or false:
        \begin{enumerate}
            \item A linear system cannot have more than three variables.
            \item A linear system in two variables and consisting of 4 equations can be consistent.
            \item Inverting the sign on each term of a linear equation is a valid row operation.
            \item $(0,0)$ is always a solution of a two-variable homogeneous linear system.
        \end{enumerate}
        
        \item Is the following system of equations a linear system? Why or why not?
        $$\begin{cases}
            3x^2 + 2y^2 = 4 \\
            4x + 3y = -1
        \end{cases}$$

        \item Suppose a three-variable linear homogeneous system is known to have additional solution(s) beyond $(0,0,0)$. How
        many solutions to the system will there be?
    \end{questions}


\end{document}