\documentclass[11pt]{exam}
\usepackage{../commonheader}
\usepackage{graphicx}

\title{Midterm Exam Review}
\date{Week 4, Session 1}

\begin{document}
\maketitle

\section{Exam Practice}
    
    \vspace{20px}
    \subsection{Linear Systems and Matrices}
    \begin{questions}
        \item Suppose a linear system has one solution. Is it consistent or inconsistent?
        \item Are the linear systems represented by the following augmented matrices consistent or inconsistent?
        $$\begin{amatrix}{2} 2 & 1 & 3 \\ 0 & 1 & 3\end{amatrix},
          \begin{amatrix}{3} 2 & -1 & 5 & 5 \\ 0 & 1 & 3 & 5 \\ 0 & 0 & 0 & 1 \end{amatrix}$$
        \item For what value of $k$ will the linear system below be consistent?
        $$\begin{cases} x - 6y = -5 \\ x - 4y = k \\ 3x - 9y = -3 \end{cases}$$
        \item If $A = \begin{bmatrix} x & 4 \\ -5 & y \end{bmatrix}$ and $A^2 = \begin{bmatrix} 29 & 24 \\ -30 & -19 \end{bmatrix}$,
        what are the values of $x$ and $y$?
        \item Reduce the following matrices to row-echelon form:
        $$\begin{bmatrix} 1 & -5 & 3 \\ -4 & 2 & -3 \end{bmatrix},
          \begin{bmatrix} 3 & 4 & 2 \\ 6 & 8 & 4 \\ 6 & 3 & 2 \end{bmatrix}$$
    \end{questions}

    \vspace{20px}
    \subsection{Homoegeneous Systems}
    \begin{questions}
        \item Build a $2 \times 2$, non-zero coefficient matrix such that the vector $\begin{bmatrix}2 \\ 4\end{bmatrix}$ is a solution
        to the corresponding homogeneous system.
        \item Suppose the coefficient matrix for an 11-variable homogeneous system has rank 10. How many solutions will the homogeneous
        system have?
    \end{questions}

    \pagebreak
    \subsection{Inverses, Transposes, and LU Decomposition}
    \begin{questions}
        \item Find the inverse (if one exists) for the following matrices:
        $$\begin{bmatrix} 1 & -3 \\ 5 & -14 \end{bmatrix}, \begin{bmatrix} 2 & 0 \\ 5 & 0 \end{bmatrix}, \begin{bmatrix} 7 & 2 \\ -11 & -3 \end{bmatrix}$$
        \item Find the tranpositions of the following matrices:
        $$\begin{bmatrix} 2 & 4 \\ 3 & 1 \end{bmatrix}, \begin{bmatrix} -2 & 0 & -3 \\ 4 & 1 & -1 \\ 1 & 0 & 1 \end{bmatrix}$$
        \item Are the following matrices symmetric? Skew symmetric?
        $$\begin{bmatrix} 1 & 2 \\ 2 & 1 \end{bmatrix},
          \begin{bmatrix} 0 & 2 \\ -2 & 0 \end{bmatrix},
          \begin{bmatrix} 1 & 2 & 3 \\ 2 & 4 & 5 \\ 3 & 5 & 3 \end{bmatrix}$$
        \item Find an LU decomposition for the following matrices:
        $$\begin{bmatrix} -6 & 9 \\ -7 & 9 \end{bmatrix}, \begin{bmatrix} 10 & -4 \\ -4 & 6 \end{bmatrix}$$
    \end{questions}

    \vspace{20px}
    \subsection{Determinants}
    \begin{questions}
        \item Find the determinants of the following $2 \times 2$ matrices:
        $$\begin{bmatrix} 4 & 3 \\ -3 & 3 \end{bmatrix}, \begin{bmatrix} 2 & 1 \\ 3 & -3 \end{bmatrix}, \begin{bmatrix} 2 & 3 \\ 0 & -1 \end{bmatrix}$$
        \item Find the determinants of the following $3 \times 3$ matrices:
        $$\begin{bmatrix} 1 & 2 & 0 \\ 3 & 5 & -2 \\ 4 & 2 & 0 \end{bmatrix}, \begin{bmatrix} 1 & 2 & 0 \\ 0 & 5 & 4 \\ 0 & 0 & -2 \end{bmatrix}$$
    \end{questions}

    \pagebreak
    \subsection{Cramer's Rule}
    We'll practice this more in the second session of this week, but it's something that may show up on your exam, so let's practice it!

    Suppose we have a linear system of the form $A\vec{x} = \vec{b}$, where $A$ is a square invertible matrix.
    Then $\vec{x} = \begin{bmatrix} x_1 & x_2 & \dots & x_n \end{bmatrix}^T$ and $x_i = \dfrac{det(A_i)}{det(A)}$, where $A_i$ is the matrix
    formed by replacing the $i$th column of $A$ with $\vec{b}$.

    As an example, consider the system defined by $A = \begin{bmatrix} -1 & -3 \\ 3 & 4 \end{bmatrix},
    \vec{b} = \begin{bmatrix} 11 \\ -13 \end{bmatrix}$.

    Then $A_1 = \begin{bmatrix} 11 & -3 \\ -13 & 4 \end{bmatrix}$ ($A$ but with its first column replaced by $\vec{b}$), \\
    and $A_2 = \begin{bmatrix} -1 & 11 \\ 3 & -13 \end{bmatrix}$ ($A$ but with its second column replaced by $\vec{b}$).

    Then $x_1 = \dfrac{det(A_1)}{det(A)}$ and $x_2 = \dfrac{det(A_2)}{det(A)}$. What are these determinants, and subsequently the solution to the
    system, $\vec{x}$?

    \vspace{20px}
    \begin{questions}
        \item Use Cramer's rule to solve the following system:
        $$\begin{cases}
            4x + 5y = 23 \\
            -x + 2y = 4
        \end{cases}$$
    \end{questions}

    \vspace{20px}
    \subsection{Properties of Determinants}
    \begin{questions}
        \item Suppose $A$ and $B$ are invertible $2 \times 2$ matrices such that $det(A) = 2$ and $det(B) = 3$. Then find the following determinants:
        \begin{enumerate}[a.]
            \item $det(AB)$
            \item $det(A^T)$
            \item $det(A^{-1})$
            \item $det(2A)$
            \item $det(B^{-1}A^{-1})$
        \end{enumerate}
    \end{questions}


\end{document}